Forecasting Gamma-Ray Bursts using Gravitational Waves

### 2018-08-27 Paper/Manuscript.tex
Sesana corrections finished.

\begin{abstract}
We explore the intriguing possibility of employing future ground-based gravitational-wave interferometers to detect the inspiral of binary neutron stars sufficiently
early to alert electromagnetic observatories so that a gamma-ray burst (GRB) can be observed in its entirety from its very beginning.
We quantify the ability to predict a GRB by computing the time a binary neutron star (BNS) system takes to inspiral from its moment of detection to its final merger. We define the moment of detection to be the instant at which the interferometer network accumulates a signal-to-noise ratio of 15. %for the BNS inspiral.
For our computations, %of advance warning times, 
we specifically consider BNS systems at luminosity distances of (i) $D\le200\,$Mpc for the three-interferometer Advanced-LIGO-Virgo network of 2020, and (ii) $D \le 1000\,$Mpc for Einstein Telescope's B and C configurations. 
In the case of Advanced LIGO-Virgo we find that we may at best get a few minutes of warning time, thus we expect no forecast of GRBs in the 2020s. 
On the other hand, Einstein Telescope will provide us with advance warning times of more than five hours for $D \le 100\,$Mpc.
Taking one hour as a benchmark advance warning time, we obtain a corresponding horizon
distance of roughly 600 Mpc for the Einstein Telescope C configuration.
%, more specifically, greater than an hour
%for BNS inspirals closer than roughly 600\,Mpc. Taking this luminosity distance to be our horizon distance, 
Using current BNS merger event rates within this volume, we show that Einstein C will forecast $\gtrsim \ord(10^2)$ GRBs in the 2030s. %with Einstein C. %with the C configuration of Einstein Telescope.
We reapply our warning-time computation to binary black hole - neutron star inspirals and find that we expect 1 to 3 tidal disruption events 
to be forecast by the same detector.
This article is intended as a pedagogical introduction to gravitational-wave astronomy written at a level accessible to
Ph.D. students, advanced undergraduates, and colleagues in astronomy and/or astrophysics who wish to learn more about the underlying physics.
Though many of our results may be known to the experts, they might nonetheless find this article motivating and exciting.
\end{abstract}
\maketitle

@@

Here, we are interested  in exploring the predictive power of near-future ground-based GW detectors. 
%thus leave the task of reducing the latency to the experts.
More specifically, we wish to find out how much early warning future GW detectors can provide us before the merger/GRB.
Ref.~\cite{Cannon:2011vi} approached this from an algorithmic perspective by developing computationally inexpensive methods to detect inspiral
signals in GW data to provide early-warning triggers \emph{before} the merger (also see Ref.~\cite{Knowles:2018hqq} for a very recent improvement to what is currently being
used in the ALV pipeline).
We instead focus on the forecasting capabilities of the ground-based interferometers at their \emph{full design} sensitivity by computing the time interval
between the merger and the instant of detection defined in terms of a certain detection criterion which we describe below.
To this end, we consider the ALV network of 2020s composed of three L-shaped interferometers operating at their design sensitivities
and Einstein Telescope of 2030s made up of V-shaped interferometers.
Ref.~\cite{Nissanke:2012dj} investigated the capabilities of three to five-interferometer detector networks of 2020s in terms of their horizon distances
to inspiralling BNSs and their sky-localization of these sources. Our work here is somewhat complementary in the sense that we are mostly concerned with
how early the GW networks can detect the inspiral before the merger.

@@

%Let us emphasize that this article is intended as a pedagogical introduction to gravitational-wave astronomy written at a level accessible to
%Ph.D. students, advanced undergraduates, and colleagues in astronomy and/or astrophysics who wish to learn more about the underlying physics.
%Be that as it may, even the experts might find our results motivating and exciting.
%
This article is organized as follows. Sec.~\ref{sec:BNS_inspiral} introduces the formulation for the leading-order evolution of the binary inspiral.
Sec.~\ref{sec:IFO_response} details the interferometer response to GWs based on interferometer topology.
Sec.~\ref{sec:idealizations} lists the various idealizations we employ to simplify our treatment and supplies justification for each.
Sec.~\ref{sec:GW170817} tests our model network and evolution using the parameters of GW170817 and the corresponding observations.
Sec.~\ref{sec:results} contains our main results presented as advance warning times in Tables~\ref{table:LIGO2020} and \ref{table:ET}; 
and as horizon distances and event rates in Table~\ref{table:horizon}.
In Sec.~\ref{sec:BH_NS}, we recompute the advance warning times in the case of binary black hole - neutron star inspirals to see whether or not 
future detectors can forecast tidal disruption events, for which we summarize our findings in Table~\ref{table:ET_BH_NS}.

@@

Another useful relation is how the inspiral time scales with respect to the orbital radius corresponding to the observed GW frequency.
From Kepler's third law, we immediately have $\dot{r}/r= -2\dot{\Omega}/(3\Omega)=-2\dot{f}/(3f)$, which, via Eq.~(\ref{eq:tau_insp}),
yields $\dot{r}/r= -1/(4\tau)$ which integrates to
%
\be
r(\tau)=r_i \left(\f{\tau}{\tau_i}\right)^{1/4}\label{eq:r_of_tau},
\ee
%
where $r_i, \tau_i=t_\text{coal}-t_i$ are the initial radius and time that the BNS
is ``picked up'' by a detector. Solving $\tau(f_i)=\tau_i$ for $f_i=\Omega_i/\pi$ using Eq.~(\ref{eq:tau_insp}) and rewriting $\Omega_i$ as a function of $r_i$ via
Kepler's third law gives us
%
\be
\tau_i = \f{5}{256}\, \f{c^5 r_i^4}{G^3 M^2 \mu}\simeq 20.37\,\text{hours}\, \left(\f{T_K}{1\,\text{s}}\right)^{8/3} \left(\f{2.8M_\odot}{M}\right)^{2/3}\,\left(\f{0.7M_\odot}{\mu}\right) \label{eq:tau_of_r},
\ee
%
where the seemingly arbitrary $0.7M_\odot$ is the value of $\mu$ for $m_1=m_2=1.4M_\odot$.

@@

\section{Interferometer response to gravitational waves}\label{sec:IFO_response}
\subsection{LIGO-Virgo configuration: L-shaped topology}\label{sec:LIGO_topo}
We start by first considering the L-shaped interferometer topology dating back to Michelson \cite{Michelson:1887zz}.
Although GWs are described by propagating tensor modes, an interferometer (IFO) can only measure a scalar quantity
known as the response function (or GW strain) which is a linear combination of the polarizations given by
%
\be
h(t) = F_+(\theta,\phi,\psi) \,h_+(t)+F_\times(\theta,\phi,\psi)\, h_\times(t) , \label{eq:detector_strain}
\ee
%
where
%%
%
\begin{align}
 F_+ &=\f{1}{2}\left(1+\cos^2\theta\right)\cos2\phi\cos 2\psi-\cos\theta \sin2\phi \sin2\psi \label{eq:F_plus},\\
 F_\times &=\f{1}{2}\left(1+\cos^2\theta\right)\cos2\phi\sin2\psi+\cos\theta \sin2\phi \cos2\psi \label{eq:F_cross}
\end{align}
%
%%
%$F_+, F_\times$ 
are the antenna pattern functions of the detectors; $\theta,\phi$ are the sky coordinates; and $\psi$ is
the angle of orientation of a detector's frame with respect to the source frame. %(there is no need for $\psi$ in the case of a single detector).
%In principle, $\psi$ can be aligned with the polarization of the incoming GWs 
%
We can not apriori know $\theta, \phi,\psi$, however, we can extract their values with some limited precision over the course of an inspiral.
The longer the inspiral stays in the detector bandwidth, the better we can estimate these angles.
%Thus, BNS inspirals which would sweep across the detection band over thousands of cycles are preferable. 
In the the case of GW170817 which swept the ALV network over roughly 3000 GW cycles, 
the initial source sky localization was 31\,deg$^2$ (later reduced to 28\,deg$^2$) using data from three IFOs \cite{GW170817}. 
For our estimations, we employ suitable averages of these angles as explained below.
For details on how to obtain Eqs.~(\ref{eq:F_plus}, \ref{eq:F_cross}) see, e.g., Sec.~4.2.1 of Ref.~\cite{SchutzLRR}.

@@

Including the error bars, the known range for neutron star masses runs from $\approx 0.5 M_\odot$ to $ \approx 3 M_\odot$ with typical values between $\approx 1 M_\odot$ and $\approx 2 M_\odot$ \cite{Ozel_Freire} (one exception being J1748-2021B with mass $2.74^{+0.21}_{-0.21} M_\odot $ \cite{Freire_et_al_2007}).

@@

 as A-LIGO's horizon will be slightly over 200\,Mpc and A-Virgo's around 130\,Mpc.

@@

%
\be
\bar\rho_{\text{ET}}(f_0,f)\equiv \f{6}{5}A h_0\, (1+z)^{-1/6}\left[\int_{f_0}^{f} d f'\, \f{f'^{-7/3}}{S^\text{ET}_n(f')}\right]^{1/2} \label{eq:ET_SNRv2},
\ee
where $f_0$ is the GW frequency at which the emitted GWs enter ET's detection band, i.e, $f_0 < f_\text{ISCO}$ such that
%
 \be
 \sqrt{S^\text{ET}_\text{n}(f_0)} = 2\sqrt{f_0}\, \tilde{H}_\text{ET}(f_0) \label{eq:f0_ET}\,
 \ee
 %
with $\tilde{H}_\text{ET}(f)$ given by the redshifted version of Eq.~(\ref{eq:H_delta})
%
\be
\tilde{H}_\text{ET}(f) = \f{1}{2}\sqrt{\f{3}{10}}\, \f{c}{D}\f{1}{(1+z)}\left(\f{G \mathcal{M}_c}{c^3}\right)^{5/6} f^{-7/6}\, , \label{eq:H_delta_redshifted}
\ee
%
where recall $\mathcal{M}_c = (1+z)M_c$.
For the detector noise $\sqrt{S^\text{ET}_n(f)}$, we adopt the ASD for the B and C configurations known as ET-B and ET-C, respectively \cite{Hild:2010id}, which we plot as the solid dark (red) and light (brown) curves in Fig.~\ref{fig:ETB2030}.
Sweeping across these are four BNS inspirals at $D=100,200,400,1000\,$Mpc represented by solid (black), dotted (blue), dashed (green),
and dot-dashed lines (gray), respectively. These strains sit much higher than the detector noise
compared with those in Fig.~\ref{fig:LIGO2020} which foretells us that the SNRs accumulated in the ET band will be much higher, 
thus yielding much longer advance warning times as we show in Table~\ref{table:ET}.
%As shown, Einstein Telescope is sensitive enough to accumulate more SNR in this $f>f_\text{ISCO}$ regime.

@@

\begin{table}[h]
\centering
\begin{tabular}{lccccccc}
%toprule
\hline\hline
$D\,$(Mpc) & \multicolumn{3}{c}{ET-B} &  & \multicolumn{3}{c}{ET-C}\T\B\\
\hline
{}& $\bar{f}_\text{ET}\,$(Hz) & \ \hspace{7mm} $T_\text{AW}$ \ \hspace{7mm} & $\bar{\rho}_{F}$ &{} & $\bar{f}_\text{ET}\,$(Hz) & \ \hspace{5mm} $T_\text{AW}\hspace{12mm}$& $\bar{\rho}_{F}$\T\B\\
%midrule
100 & $\approx\,$6.72 &  47.0\,minutes & 306 &{\qquad} & $\approx\,$3.27 & 5.34\,hours\ & 365\T\\
200 & $\approx\,$11.2 & 11.6\,minutes & 152 &{\qquad} & $\approx\,$4.10 & 2.87\,hours\ & 182 \\
400 & $\approx\,$18.2 & 3.00\,minutes & 75.7 &{\qquad} & $\approx\,$5.06 & 1.51\,hours\ & 90.5\\
1000 & $\approx\,$41.3 &17.2\,seconds & 29.8& \qquad &    $\approx\,$6.76 & 35.6\,minutes & 35.6  \\
% OLD version
%100 & 6.7237 &  2818 & 306.0 &{\qquad} & 3.2692 & 19212 & 365.4\T\\
%200 & 11.2104 & 698.6 & 152.4 &{\qquad} & 4.1028 & 10138 & 182.1 \\
%400 & 18.2234 & 179.8 & 75.74 &{\qquad} & 5.0607 & 5437 & 90.48\\
%1000 & 41.3537 &17.2 & 29.79& \qquad & 6.7553 & 2138 & 35.60  \\
\hline\hline
%bottomrule
\end{tabular}
\caption{Forecasting capabilities of Einstein Telescope. 
%We can expect over \emph{five} hours of early warning for nearby sources. 
ET-B and ET-C refer to the different configurations shown in Fig.~\ref{fig:ETB2030}. For the advance warning times, we only present the result of the more accurate 3.5PN computation. $\bar{f}_\text{ET}$ is the threshold frequency at which
ET-B/C accumulate SNR of 15 which we take to be our detection criterion. Note that both $T_\text{AW}$ and $\bar\rho_F$ are larger for ET-C
due to its improved sensitivity in the $1\,\text{Hz}\lesssim f\lesssim 30\,$Hz regime compared to ET-B as is clear in Fig.~\ref{fig:ETB2030}.
These results and those of Table.~\ref{table:LIGO2020} are summarized in Fig.~\ref{fig:summary}.}\label{table:ET}
\end{table}
%
%
As the results of Table~\ref{table:ET} indicate, ET-B will provide $\gtrsim 10\,$ minutes of warning time for BNS inspirals at $D\lesssim 200\,$Mpc.
Instead of speculating whether or not this is enough to forecast GRBs, we turn our attention to ET-C's forecasting capabilities.
As can be seen from Fig.~\ref{fig:ETB2030}, ET-C's increased sensitivity in the $1\,\text{Hz}\lesssim f\lesssim 30\,$Hz regime will significantly lengthen the advance warning times.
This is best highlighted by the sixth colum of Table~\ref{table:ET}, 
where we show that ET-C will provide up to \emph{five hours} of early warning time for nearby ($D\lesssim 100\,$Mpc) neutron star mergers.
Moreover, BNS inspirals as far as 200\,Mpc (the ALV network's horizon distance) will persist in the ET-C sensitivity band for a few hours after their initial detection ($T_\text{AW} > 2\,$ hours).
Current BNS merger rate, inferred from A-LIGO's O1-O2 periods, implies that $\sim 40$ to $\sim600$ sources within a volume of $(200\,\text{Mpc})^3$ will be detected yearly by ET-C.

Given Moore's law and new search approaches (e.g., see Ref.~\cite{Gabbard:2017lja} for one based on deep learning) 
we speculate that the advances in the detection, localization and early-warning algorithms in the next decade will reduce the required advance
warning time to less than an hour; thus we set $T_\text{AW} = 1\,\text{hour}$ and obtain corresponding BNS horizon distances of $D_H=87$ and $613\,$Mpc ($z\approx 0.02,0.127$) for ET-B and ET-C, respectively.
In other words, all inspiralling BNSs closer than $D_H$ will give us more than an hour's leeway before the GRB (employing the detection criterion $\text{SNR} =15$ as before).
The volume set by $D=D_H$ translates to a BNS merger rate of $\sim\mathcal{O}(1)\,\text{yr}^{-1}$ for ET-B and
$\approx 355^{+730}_{-280}\,\text{yr}^{-1}$ for ET-C. Additionally, each source within $D_H$ will accumulate $\bar\rho_F \gtrsim 420, 58$ over its inspiral
in the ET-B, ET-C bands, respectively. These results are summarized in Table~\ref{table:horizon}.

>>>>

\begin{abstract}
We explore the intriguing possibility of employing future ground-based gravitational-wave interferometers to detect the inspiral of binary neutron stars sufficiently
early to alert electromagnetic observatories so that a gamma-ray burst (GRB) can be observed in its entirety from its very beginning.
We quantify the ability to predict a GRB by computing the time a binary neutron star (BNS) system takes to inspiral from its moment of detection to its final merger. We define the moment of detection to be the instant at which the interferometer network accumulates a signal-to-noise ratio of 15. %for the BNS inspiral.
For our computations, %of advance warning times, 
we specifically consider BNS systems at luminosity distances of (i) $D\le200\,$Mpc for the three-interferometer Advanced-LIGO-Virgo network of 2020, and (ii) $D \le 1000\,$Mpc for Einstein Telescope's B and C configurations. 
In the case of Advanced LIGO-Virgo we find that we may at best get a few minutes of warning time, thus we expect no forecast of GRBs in the 2020s. 
On the other hand, Einstein Telescope will provide us with advance warning times of more than five hours for $D \le 100\,$Mpc.
Taking one hour as a benchmark advance warning time, we obtain a corresponding range of roughly 600 Mpc for the Einstein Telescope C configuration.
%, more specifically, greater than an hour
%for BNS inspirals closer than roughly 600\,Mpc. Taking this luminosity distance to be our horizon distance, 
Using current BNS merger event rates within this volume, we show that Einstein C will forecast $\gtrsim \ord(10^2)$ GRBs in the 2030s. %with Einstein C. %with the C configuration of Einstein Telescope.
We reapply our warning-time computation to binary black hole - neutron star inspirals and find that we expect 1 to 3 tidal disruption events 
to be forecast by the same detector.
This article is intended as a pedagogical introduction to gravitational-wave astronomy written at a level accessible to
Ph.D. students, advanced undergraduates, and colleagues in astronomy and/or astrophysics who wish to learn more about the underlying physics.
Though many of our results may be known to the experts, they might nonetheless find this article motivating and exciting.
\end{abstract}
\maketitle

@@

Here, we are interested  in exploring the predictive power of near-future ground-based GW detectors. 
%thus leave the task of reducing the latency to the experts.
More specifically, we wish to find out how much early warning future GW detectors can provide us before the merger/GRB.
Ref.~\cite{Cannon:2011vi} approached this from an algorithmic perspective by developing computationally inexpensive methods to detect inspiral
signals in GW data to provide early-warning triggers \emph{before} the merger (also see Ref.~\cite{Knowles:2018hqq} for a very recent improvement to what is currently being
used in the ALV pipeline).
We instead focus on the forecasting capabilities of the ground-based interferometers at their \emph{full design} sensitivity by computing the time interval
between the merger and the instant of detection defined in terms of a certain detection criterion which we describe below.
To this end, we consider the ALV network of 2020s composed of three L-shaped interferometers operating at their design sensitivities
and Einstein Telescope of 2030s made up of V-shaped interferometers.
Ref.~\cite{Nissanke:2012dj} investigated the capabilities of three to five-interferometer detector networks of 2020s in terms of their ranges
to inspiralling BNSs and their sky-localization of these sources. Our work here is somewhat complementary in the sense that we are mostly concerned with
how early the GW networks can detect the inspiral before the merger.

@@

%Let us emphasize that this article is intended as a pedagogical introduction to gravitational-wave astronomy written at a level accessible to
%Ph.D. students, advanced undergraduates, and colleagues in astronomy and/or astrophysics who wish to learn more about the underlying physics.
%Be that as it may, even the experts might find our results motivating and exciting.
%
This article is organized as follows. Sec.~\ref{sec:BNS_inspiral} introduces the formulation for the leading-order evolution of the binary inspiral.
Sec.~\ref{sec:IFO_response} details the interferometer response to GWs based on interferometer topology.
Sec.~\ref{sec:idealizations} lists the various idealizations we employ to simplify our treatment and supplies justification for each.
Sec.~\ref{sec:GW170817} tests our model network and evolution using the parameters of GW170817 and the corresponding observations.
Sec.~\ref{sec:results} contains our main results presented as advance warning times in Tables~\ref{table:LIGO2020} and \ref{table:ET}; 
and as ranges and event rates in Table~\ref{table:horizon}.
In Sec.~\ref{sec:BH_NS}, we recompute the advance warning times in the case of binary black hole - neutron star inspirals to see whether or not 
future detectors can forecast tidal disruption events, for which we summarize our findings in Table~\ref{table:ET_BH_NS}.

@@

Another useful relation is how the inspiral time scales with respect to the orbital radius corresponding to the observed GW frequency.
From Kepler's third law, we immediately have $\dot{r}/r= -2\dot{\Omega}/(3\Omega)=-2\dot{f}/(3f)$, which, via Eq.~(\ref{eq:tau_insp}),
yields $\dot{r}/r= -1/(4\tau)$ which integrates to
%
\be
r(\tau)=r_i \left(\f{\tau}{\tau_i}\right)^{1/4}\label{eq:r_of_tau},
\ee
%
where $r_i, \tau_i=t_\text{coal}-t_i$ are the initial radius and time that the BNS
is ``picked up'' by a detector. Solving $\tau(f_i)=\tau_i$ for $f_i=\Omega_i/\pi$ using Eq.~(\ref{eq:tau_insp}) and rewriting $\Omega_i$ as a function of $r_i$ via
Kepler's third law gives us
%
\be
\tau_i = \f{5}{256}\, \f{c^5 r_i^4}{G^3 M^2 \mu}\simeq 1.024\,\text{hr} \left(\f{r_i}{1000\,\text{km}}\right)^4 \left(\f{2.8M_\odot}{M}\right)^2\,\left(\f{0.7M_\odot}{\mu}\right)
%\simeq 20.37\,\text{hours}\, \left(\f{T_K}{1\,\text{s}}\right)^{8/3} \left(\f{2.8M_\odot}{M}\right)^{2/3}\,\left(\f{0.7M_\odot}{\mu}\right)
\label{eq:tau_of_r},
\ee
%
where the seemingly arbitrary $0.7M_\odot$ is the value of $\mu$ for $m_1=m_2=1.4M_\odot$.

@@

\section{Interferometer response to gravitational waves}\label{sec:IFO_response}
\subsection{LIGO-Virgo configuration: L-shaped topology}\label{sec:LIGO_topo}
We start by first considering the L-shaped interferometer topology dating back to Michelson \cite{Michelson:1887zz}.
Although GWs are described by propagating tensor modes, an interferometer (IFO) can only measure a scalar quantity
known as the response function (or GW strain) which is a linear combination of the polarizations given by
%
\be
h(t) = F_+(\theta,\phi,\psi) \,h_+(t)+F_\times(\theta,\phi,\psi)\, h_\times(t) , \label{eq:detector_strain}
\ee
%
where
%%
%
\begin{align}
 F_+ &=\f{1}{2}\left(1+\cos^2\theta\right)\cos2\phi\cos 2\psi-\cos\theta \sin2\phi \sin2\psi \label{eq:F_plus},\\
 F_\times &=\f{1}{2}\left(1+\cos^2\theta\right)\cos2\phi\sin2\psi+\cos\theta \sin2\phi \cos2\psi \label{eq:F_cross}
\end{align}
%
%%
%$F_+, F_\times$ 
are the antenna pattern functions of the detectors; $\theta,\phi$ are the sky coordinates; and $\psi$ is
the angle of orientation of a detector's frame with respect to the source frame. %(there is no need for $\psi$ in the case of a single detector).
%In principle, $\psi$ can be aligned with the polarization of the incoming GWs 
%
We can not apriori know $\theta, \phi,\psi$, however, we can extract their values with some limited precision over the course of an inspiral.
The longer the inspiral stays in the detector bandwidth, the better we can estimate these angles.
%Thus, BNS inspirals which would sweep across the detection band over thousands of cycles are preferable. 
In the the case of GW170817 which swept the ALV network over roughly 3000 GW cycles, 
the initial rapid source sky localization was 31\,deg$^2$ (later reduced to 28\,deg$^2$) using data from three IFOs \cite{GW170817}. 
For our estimations, we employ suitable averages of these angles as explained below.
For details on how to obtain Eqs.~(\ref{eq:F_plus}, \ref{eq:F_cross}) see, e.g., Sec.~4.2.1 of Ref.~\cite{SchutzLRR}.

@@

Including the error bars, the known range for neutron star masses runs from $\approx 0.5 M_\odot$ to $ \approx 3 M_\odot$ with typical values between $\approx 1 M_\odot$ and $\approx 2 M_\odot$ \cite{Ozel_Freire} (one exception being J1748-2021B with mass $2.74^{+0.21}_{-0.21} M_\odot $ \cite{Freire_et_al_2007}, but this measurement is not secure \cite{Lattimer:2012nd}\footnote{We thank Dr. Alberto Sesana of University of Birmingham for pointing this out.}).

@@

 as A-LIGO's range will be slightly over 200\,Mpc and A-Virgo's around 130\,Mpc.

@@

%
\be
\bar\rho_{\text{ET}}(f_0,f)\equiv \f{6}{5}A h_0\, (1+z)^{-1/6}\left[\int_{f_0}^{f} d f'\, \f{f'^{-7/3}}{S^\text{ET}_n(f')}\right]^{1/2} \label{eq:ET_SNRv2},
\ee
where $f_0$ is the GW frequency at which the emitted GWs enter ET's detection band, i.e, $f_0 < f_\text{ISCO}$ such that
%
 \be
 \sqrt{S^\text{ET}_\text{n}(f_0)} = 2\sqrt{f_0}\, \tilde{H}_\text{ET}(f_0) \label{eq:f0_ET}\,
 \ee
 %
with $\tilde{H}_\text{ET}(f)$ given by the redshifted version of Eq.~(\ref{eq:H_delta})
%
\be
\tilde{H}_\text{ET}(f) = \f{1}{2}\sqrt{\f{3}{10}}\,\pi^{-2/3} \f{c}{D}\f{1}{(1+z)}\left(\f{G \mathcal{M}_c}{c^3}\right)^{5/6} f^{-7/6}\, , \label{eq:H_delta_redshifted}
\ee
%
where recall $\mathcal{M}_c = (1+z)M_c$.
For the detector noise $\sqrt{S^\text{ET}_n(f)}$, we adopt the ASD for the B and C configurations known as ET-B and ET-C, respectively \cite{Hild:2010id}, which we plot as the solid dark (red) and light (brown) curves in Fig.~\ref{fig:ETB2030}.
Sweeping across these are four BNS inspirals at $D=100,200,400,1000\,$Mpc represented by solid (black), dotted (blue), dashed (green),
and dot-dashed lines (gray), respectively. These strains sit much higher than the detector noise
compared with those in Fig.~\ref{fig:LIGO2020} which foretells us that the SNRs accumulated in the ET band will be much higher, 
thus yielding much longer advance warning times as we show in Table~\ref{table:ET}.
%As shown, Einstein Telescope is sensitive enough to accumulate more SNR in this $f>f_\text{ISCO}$ regime.

@@

\begin{table}[h]
\centering
\begin{tabular}{lccccccc}
%toprule
\hline\hline
$D\,$(Mpc) & \multicolumn{3}{c}{ET-B} &  & \multicolumn{3}{c}{ET-C}\T\B\\
\hline
{}& $\bar{f}_\text{ET}\,$(Hz) & \ \hspace{7mm} $T_\text{AW}$ \ \hspace{7mm} & $\bar{\rho}_{F}$ &{} & $\bar{f}_\text{ET}\,$(Hz) & \ \hspace{5mm} $T_\text{AW}\hspace{12mm}$& $\bar{\rho}_{F}$\T\B\\
%midrule
100 & $\approx\,$6.72 &  47.0\,minutes & 306 &{\qquad} & $\approx\,$3.27 & 5.34\,hours\ & 365\T\\
200 & $\approx\,$11.2 & 11.6\,minutes & 152 &{\qquad} & $\approx\,$4.10 & 2.87\,hours\ & 182 \\
400 & $\approx\,$18.2 & 3.00\,minutes & 75.7 &{\qquad} & $\approx\,$5.06 & 1.51\,hours\ & 90.5\\
1000 & $\approx\,$41.3 &17.2\,seconds & 29.8& \qquad &    $\approx\,$6.76 & 35.6\,minutes & 35.6  \\
% OLD version
%100 & 6.7237 &  2818 & 306.0 &{\qquad} & 3.2692 & 19212 & 365.4\T\\
%200 & 11.2104 & 698.6 & 152.4 &{\qquad} & 4.1028 & 10138 & 182.1 \\
%400 & 18.2234 & 179.8 & 75.74 &{\qquad} & 5.0607 & 5437 & 90.48\\
%1000 & 41.3537 &17.2 & 29.79& \qquad & 6.7553 & 2138 & 35.60  \\
\hline\hline
%bottomrule
\end{tabular}
\caption{Forecasting capabilities of Einstein Telescope. 
%We can expect over \emph{five} hours of early warning for nearby sources. 
ET-B and ET-C refer to the different configurations shown in Fig.~\ref{fig:ETB2030}. For the advance warning times, we only present the result of the more accurate 3.5PN computation. $\bar{f}_\text{ET}$ is the threshold frequency at which
ET-B/C accumulate SNR of 15 which we take to be our detection criterion. Note that both $T_\text{AW}$ and $\bar\rho_F$ are larger for ET-C
due to its improved sensitivity in the $1\,\text{Hz}\lesssim f\lesssim 30\,$Hz regime compared to ET-B as is clear in Fig.~\ref{fig:ETB2030}.
These results and those of Table.~\ref{table:LIGO2020} are summarized in Fig.~\ref{fig:summary}.}\label{table:ET}
\end{table}
%
%
As the results of Table~\ref{table:ET} indicate, ET-B will provide $\gtrsim 10\,$ minutes of warning time for BNS inspirals at $D\lesssim 200\,$Mpc.
Instead of speculating whether or not this is enough to forecast GRBs, we turn our attention to ET-C's forecasting capabilities.
As can be seen from Fig.~\ref{fig:ETB2030}, ET-C's increased sensitivity in the $1\,\text{Hz}\lesssim f\lesssim 30\,$Hz regime will significantly lengthen the advance warning times.
This is best highlighted by the sixth colum of Table~\ref{table:ET}, 
where we show that ET-C will provide up to \emph{five hours} of early warning time for nearby ($D\lesssim 100\,$Mpc) neutron star mergers.
Moreover, BNS inspirals as far as 200\,Mpc (the ALV network's range) will persist in the ET-C sensitivity band for a few hours after their initial detection ($T_\text{AW} > 2\,$ hours).
Current BNS merger rate, inferred from A-LIGO's O1-O2 periods, implies that $\sim 40$ to $\sim600$ sources within a volume of $(200\,\text{Mpc})^3$ will be detected yearly by ET-C.

Given Moore's law and new search approaches (e.g., see Ref.~\cite{Gabbard:2017lja} for one based on deep learning) 
we speculate that the advances in the detection, localization and early-warning algorithms in the next decade will reduce the required advance
warning time to less than an hour; thus we set $T_\text{AW} = 1\,\text{hour}$ and obtain corresponding BNS ranges of $D_\text{BNS}=87$ and $613\,$Mpc ($z\approx 0.02,0.127$) for ET-B and ET-C, respectively.
In other words, all inspiralling BNSs closer than $D_\text{BNS}$ will give us more than an hour's leeway before the GRB (employing the detection criterion $\text{SNR} =15$ as before).
The volume set by $D_\text{BNS}$ translates to a BNS merger rate of $\sim\mathcal{O}(1)\,\text{yr}^{-1}$ for ET-B and
$\approx 355^{+730}_{-280}\,\text{yr}^{-1}$ for ET-C. Additionally, each source within $D_\text{BNS}$ will accumulate $\bar\rho_F \gtrsim 420, 58$ over its inspiral
in the ET-B, ET-C bands, respectively. These results are summarized in Table~\ref{table:horizon}.

Both Tables~\ref{table:ET} and \ref{table:horizon} clearly show that ET will yield superb SNRs for sources out to 1\,Gpc. 
With such high SNRs we expect that a future network consisting of ET \emph{and} LIGO Voyager will be able to localize half the sources 
inside a radius of $D\approx 1000\,$Mpc to within $ 10\,\text{deg}^2$ and more than half the sources within $D\lesssim 400\,$Mpc to $\approx 1\,\text{deg}^2$ \cite{Mills:2017urp}.
Let us reemphasize that such good localization requires at least a second detector with comparable sensitivity to ET as ET alone has rather poor localization.
Erring on the side of optimism, we conclude that the 2030s hold in store real-time observations of dozens to hundreds of gamma-ray bursts
\footnote{As the ET construction/commissioning timeline is currently uncertain, we can only speculate that ET-C will be operational after 2030,
which is roughly in line with the milli-Hertz space interferometer LISA's timeline \cite{Audley:2017drz}.}.
%It is also reasonable to speculate that by that time the detection, localization and early-warning algorithms will have become sufficiently sophisticated
%to alert the electromagnetic observatories within an hour of the initial detection (SNR = 15). 
We summarize our main findings for both the ALV network and ET-B/C in Fig.~\ref{fig:summary}.
\begin{tabular}{lccc}
%toprule
 & ET-B & & ET-C\T\B\\
  $D_\text{BNS} $& 87\,Mpc& &{613\,Mpc}\T\\
  $R(D_\text{BNS}) $& $1^{+2}_{-1}\,\text{yr}^{-1}$&\hspace{5mm} &{$355^{+730}_{-280}\,\text{yr}^{-1}$}\T\\
  $\bar\rho_F(D_\text{BNS})$ & 420 &&{58}\T\B\\
\caption{Horizon distances of ET-B and ET-C assuming $T_\text{AW} =1\,$hour. $R(D_\text{BNS})$ is the BNS merger rate within a volume of $D_\text{BNS}^3$
obtained by rescaling the rate inferred from Advanced LIGO's O1, O2 observing periods \cite{GW170817}. $\bar\rho_F(D_\text{BNS})$ is the total SNR accumulated due to a BNS inspiralling at $D_\text{BNS}$ [see Eq.~(\ref{eq:rhoF_ET})].}\label{table:horizon}
\includegraphics[width=\linewidth]{../Figures/Main_results.pdf}
\caption{Summary of our major findings for both the Advanced-LIGO-Virgo (ALV) network circa 2020 and the Einstein Telescope's B and C configurations
ca. 2030. The light coloured (gray) data shows $T_\text{AW}$ for the ALV2020 (dotted), ET-B (dashed), and ET-C (solid), respectively, with the corresponding axis on the left.
The connecting light (gray) lines are drawn only for visualization purposes unlike the dark (blue) lines coming from the $D^{-1}$ powerlaw of the SNR [see Eq.~(\ref{eq:ET_SNRv2})] which is plotted on the right axis. 
We used $D=50,100,200\,$Mpc to obtain the ALV results (Sec.~\ref{Sec:ALIGO2020}) and $D=100,200,400,1000\,$Mpc for the ET results.}\label{fig:summary}
% LISA does not see these sources, not sensitive enough!
\section{Binary black hole neutron star systems}\label{sec:BH_NS}
In this section, we investigate the possibility of witnessing the tidal distruption of a neutron star by a black hole before the merger.
The treatment of the inspiral is the same as before, however the fate of the neutron star depends strongly on the system's intrinsic parameters
such as the black hole mass and spin, and the neutron star's equation of state (EOS) and compactness $\mathcal{C}_\text{NS}\equiv G M_\text{NS}/(R_\text{NS}c^2)$. 
The evolution of black hole-neutron star (BH-NS) binaries is a very active field of research at the interface of numerical relativity and high-energy astrophysics
(see Ref. \cite{Shibata:2011jka} for a comprehensive review) requiring fully general relativistic magnetohydrodynamic treatment including neutrino transport equations.
The complexity of evolving the dynamics of these systems require large computational resources and long runtimes.
%running on large-scale clusters.
As such, the details are beyond the scope of this article, but we can use a mostly Newtonian treatment to serve our purposes. %, however the numerical simulationshave verified
